---
title: "Process_PE_count_tsv_files"
author: "Guang"
date: "6/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## List all tsv file

There are multiple tsv files related to pair-end sam files. This script is designed to process all the tsv files and then merge the data together to a single summary tsv file.

```{r list-all-tsv-files, eval=TRUE}
all_tsv_files <- list.files(pattern = ".*different_mapping_count.tsv")
```

## Processing one tsv file to confirm tsv file content

This part will not be included into the final code. This part use one tsv file as input and check what are the contents inside it so that a function will be defined later which will be used to analyze all tsv files in batch mode(using apply() function).
```{r process-single-tsv-file, eval=TRUE}
library(data.table)
one_tsv_data <- fread(all_tsv_files[1], header = TRUE, 
                      stringsAsFactors = FALSE)
# unique(one_tsv_data$Read1_Uniqueness)
# unique(one_tsv_data$Read2_Uniqueness)
read1_uniq_right_sum <- sum(one_tsv_data$Read1_Uniqueness == "Uniq_right")
read1_uniq_right_0_score_sum <- sum(one_tsv_data$Read1_Uniqueness == "Uniq_right_0_score")
read1_nonUniq_right_sum <- sum(one_tsv_data$Read1_Uniqueness == "nonUniq_right")
read1_uniq_wrong_sum <- sum(one_tsv_data$Read1_Uniqueness == "Uniq_wrong")
read1_uniq_wrong_0_score_sum<- sum(one_tsv_data$Read1_Uniqueness == "Uniq_wrong_0_score")
read1_nonUniq_wrong_sum <- sum(one_tsv_data$Read1_Uniqueness == "nonUniq_wrong")

read2_uniq_right_sum <- sum(one_tsv_data$Read2_Uniqueness == "Uniq_right")
read2_uniq_right_0_score_sum <- sum(one_tsv_data$Read2_Uniqueness == "Uniq_right_0_score")
read2_nonUniq_right_sum <- sum(one_tsv_data$Read2_Uniqueness == "nonUniq_right")
read2_uniq_wrong_sum <- sum(one_tsv_data$Read2_Uniqueness == "Uniq_wrong")
read2_uniq_wrong_0_score_sum<- sum(one_tsv_data$Read2_Uniqueness == "Uniq_wrong_0_score")
read2_nonUniq_wrong_sum <- sum(one_tsv_data$Read2_Uniqueness == "nonUniq_wrong")

unique_read1_tags <- unique(one_tsv_data$Read1_Uniqueness)
read1_tags_sum <- rep(0, length(unique_read1_tags))
for(i in 1: length(unique_read1_tags) )
{
  read1_tags_sum[i] <- sum(one_tsv_data$Read1_Uniqueness== unique_read1_tags[i])
}
names(read1_tags_sum) <- paste0("read1_", unique_read1_tags)
read1_tags_sum

unique_read2_tags <- unique(one_tsv_data$Read2_Uniqueness)
read2_tags_sum <- rep(0, length(unique_read2_tags))
for(i in 1: length(unique_read2_tags) )
{
  read2_tags_sum[i] <- sum(one_tsv_data$Read2_Uniqueness== unique_read2_tags[i])
}
names(read2_tags_sum) <- paste0("read2_", unique_read2_tags)
read2_tags_sum

df <- data.frame(c(read1_tags_sum, read2_tags_sum))
transverse_df <- t(df)
row.names(transverse_df) <- all_tsv_files[1]
transverse_df


## Stupid but stable method:
sum_df <- data.frame(read1_uniq_right = read1_uniq_right_sum,
                     read1_uniq_right_0_score = read1_uniq_right_0_score_sum,
                     read1_nonUniq_right = read1_nonUniq_right_sum,
                     read1_uniq_wrong = read1_uniq_wrong_sum,
                     read1_uniq_wrong_0_score = read1_uniq_wrong_0_score_sum,
                     read1_nonUniq_wrong = read1_nonUniq_wrong_sum,
                      
                     read2_uniq_right = read2_uniq_right_sum,
                     read2_uniq_right_0_score = read2_uniq_right_0_score_sum,
                     read2_nonUniq_right = read2_nonUniq_right_sum,
                     read2_uniq_wrong = read2_uniq_wrong_sum,
                     read2_uniq_wrong_0_score = read2_uniq_wrong_0_score_sum,
                     read2_nonUniq_wrong = read2_nonUniq_wrong_sum)
row_name <- gsub("GRCm38.100.genome.reads.","",all_tsv_files[1])
row_name <- gsub(".aln.pe.sam.different_mapping_count.tsv","",row_name)
row_name
row.names(sum_df) <- row_name
sum_df
```



## Function to process one tsv file

The function here is used to do statistics of mapping result of one single tsv file from one aligned pair-end sam file.

```{r summary-of-one-tsv-file}
summary_count_tsv <- function(one_tsv_file)
{
  library(data.table)
  one_tsv_data <- fread(one_tsv_file, header = TRUE, stringsAsFactors = FALSE)
  
  one_tsv_data$read1_uniq_right_include_0_score <- (one_tsv_data$Read1_Uniqueness == "Uniq_right" |
                                                   one_tsv_data$Read1_Uniqueness == "Uniq_right_0_score")
  one_tsv_data$read2_uniq_right_include_0_score <- (one_tsv_data$Read2_Uniqueness == "Uniq_right" |
                                                   one_tsv_data$Read2_Uniqueness == "Uniq_right_0_score")
  
  # include 0 scored uniq_right below
  one_tsv_data$read1_uniq_right_read2_not <- (one_tsv_data$read1_uniq_right_include_0_score &
                                               (!one_tsv_data$read2_uniq_right_include_0_score) )
  one_tsv_data$read2_uniq_right_read1_not <- ( (!one_tsv_data$read1_uniq_right_include_0_score) &
                                                one_tsv_data$read2_uniq_right_include_0_score )
  one_tsv_data$read1_or_read2_uniq <- (one_tsv_data$read1_uniq_right_include_0_score | 
                                      one_tsv_data$read2_uniq_right_include_0_score)
  one_tsv_data$read1_and_read2_uniq <- (one_tsv_data$read1_uniq_right_include_0_score & 
                                      one_tsv_data$read2_uniq_right_include_0_score)
  
  read1_uniq_right_include_0_score_sum <- sum(one_tsv_data$read1_uniq_right_include_0_score)
  read2_uniq_right_include_0_score_sum <- sum(one_tsv_data$read2_uniq_right_include_0_score)
  
  read1_Uniqright_read2_not_sum <- sum(one_tsv_data$read1_uniq_right_read2_not)
  read1_not_read2_Uniqright_sum <- sum(one_tsv_data$read2_uniq_right_read1_not)
  read1_or_read2_Uniq_sum <- sum(one_tsv_data$read1_or_read2_uniq)
  read1_and_read2_Uniq_sum <- sum(one_tsv_data$read1_and_read2_uniq)
  
  read1_uniq_right_sum <- sum(one_tsv_data$Read1_Uniqueness == "Uniq_right")
  read1_uniq_right_0_score_sum <- sum(one_tsv_data$Read1_Uniqueness == "Uniq_right_0_score")
  read1_nonUniq_right_sum <- sum(one_tsv_data$Read1_Uniqueness == "nonUniq_right")
  read1_uniq_wrong_sum <- sum(one_tsv_data$Read1_Uniqueness == "Uniq_wrong")
  read1_uniq_wrong_0_score_sum<- sum(one_tsv_data$Read1_Uniqueness == "Uniq_wrong_0_score")
  read1_nonUniq_wrong_scored_sum <- sum(one_tsv_data$Read1_Uniqueness == "nonUniq_wrong_scored")
  read1_nonUniq_wrong_0_score_sum <- sum(one_tsv_data$Read1_Uniqueness == "nonUniq_wrong")
  read1_nonUniq_wrong_sum <- sum(one_tsv_data$Read1_Uniqueness == "nonUniq_wrong" |
                                 one_tsv_data$Read1_Uniqueness == "nonUniq_wrong_scored")
  
  
  
  read2_uniq_right_sum <- sum(one_tsv_data$Read2_Uniqueness == "Uniq_right")
  read2_uniq_right_0_score_sum <- sum(one_tsv_data$Read2_Uniqueness == "Uniq_right_0_score")
  read2_nonUniq_right_sum <- sum(one_tsv_data$Read2_Uniqueness == "nonUniq_right")
  read2_uniq_wrong_sum <- sum(one_tsv_data$Read2_Uniqueness == "Uniq_wrong")
  read2_uniq_wrong_0_score_sum<- sum(one_tsv_data$Read2_Uniqueness == "Uniq_wrong_0_score")
  read2_nonUniq_wrong_scored_sum <- sum(one_tsv_data$Read2_Uniqueness == "nonUniq_wrong_scored")
  read2_nonUniq_wrong_0_score_sum <- sum(one_tsv_data$Read2_Uniqueness == "nonUniq_wrong")
  read2_nonUniq_wrong_sum <- sum(one_tsv_data$Read2_Uniqueness == "nonUniq_wrong" |
                                 one_tsv_data$Read2_Uniqueness == "nonUniq_wrong_scored")
  
  
  # one row data frame: row.name is tsv file name; col.names are tags.
  ## Stupid but stable method:
  sample_name <- gsub("GRCm38.100.genome.reads.","",one_tsv_file)
  sample_name <- gsub(".aln.pe.sam.different_mapping_count.tsv","",sample_name)
  # sample_name
  sum_df <- data.frame(sample = sample_name,
                       read1_uniq_right = read1_uniq_right_sum,
                       read1_uniq_right_0_score = read1_uniq_right_0_score_sum,
                       read1_uniq_right_include_0_score = read1_uniq_right_include_0_score_sum,
                       read1_nonUniq_right = read1_nonUniq_right_sum,
                       read1_uniq_wrong = read1_uniq_wrong_sum,
                       read1_uniq_wrong_0_score = read1_uniq_wrong_0_score_sum,
                       read1_nonUniq_wrong_scored = read1_nonUniq_wrong_scored_sum, 
                       read1_nonUniq_wrong_0_score = read1_nonUniq_wrong_0_score_sum, 
                       read1_nonUniq_wrong = read1_nonUniq_wrong_sum,
                        
                       read2_uniq_right = read2_uniq_right_sum,
                       read2_uniq_right_0_score = read2_uniq_right_0_score_sum,
                       read2_uniq_right_include_0_score = read2_uniq_right_include_0_score_sum, 
                       read2_nonUniq_right = read2_nonUniq_right_sum,
                       read2_uniq_wrong = read2_uniq_wrong_sum,
                       read2_uniq_wrong_0_score = read2_uniq_wrong_0_score_sum,
                       read2_nonUniq_wrong_scored = read2_nonUniq_wrong_scored_sum,
                       read2_nonUniq_wrong_0_score = read2_nonUniq_wrong_0_score_sum, 
                       read2_nonUniq_wrong = read2_nonUniq_wrong_sum,
                       
                       # read1_uniq_right_include_0_score = read1_uniq_right_include_0_score_sum,
                       # read2_uniq_right_include_0_score = read2_uniq_right_include_0_score_sum,
                       read1_uniq_right_read2_not = read1_Uniqright_read2_not_sum,
                       read1_not_read2_uniq_right = read1_not_read2_Uniqright_sum,
                       read1_or_read2_uniq = read1_or_read2_Uniq_sum,
                       read1_and_read2_uniq = read1_and_read2_Uniq_sum)
  return(sum_df)
}
summary_count_tsv(all_tsv_files[1])
```

## Apply the summary function to all tsv files

Here I will use a `apply()` function to all tsv files so that I will get statistics of read mapping tags in each tsv file.

```{r batch-tsv-summary, eval=TRUE}
summary_data_all_tsv <- lapply(all_tsv_files, summary_count_tsv)
str(summary_data_all_tsv)
all_tsv_df <- Reduce( function(x,y) merge(x,y, all=TRUE), summary_data_all_tsv)
# all_tsv_df
write.table(all_tsv_df,
            file = "summary_of_mapping_result_all_SAMs20201007.tsv",
            col.names = TRUE,
            row.names = FALSE,
            sep = '\t')
```


